### 2014-02-18 latex-template/extended-abstract.tex

\documentclass{iitsrc}
\usepackage[utf8]{inputenc}

% Please do not remove the following command:
\editpages{1}{8}

\title{Guidelines for Authors of Extended Abstracts\\Published in Proceedings of IIT.SRC 2014}
\titlerunning{Guidelines for Authors of Papers Published in Proceedings of IIT.SRC 2014}
\author{Ján}{Novák}
\author{Martin}{Kováč}
\supervision{\bc~/~\ms~/~\phd}
            {\info~/~\compeng~/~\softeng~/~\infosys~/~\appinf}
            {Professor~/~Assoc.~Professor~/~Dr.~Xxxx Yyyyy}
            {\iai~/~\iise~/~\icsn, \fiit}
% Within the `affiliation' parameter you can use the `\fiit' macro to
% generate the full name of our faculty.
\mail{novak@fiit.stuba.sk}
\field{To Be Added by Editor}

\begin{document}

\section*{Extended Abstract}

This is a simple example of an ``extended abstract''. These documents should not exceed two pages. Please, do not use the
\begin{verbatim}
    \begin{abstract}
    ...
    \end{abstract}
\end{verbatim}
environment. Instead, put your whole text in a single unnumbered section introduced with the following command:
\begin{verbatim}
    \section*{Extended Abstract}
\end{verbatim}
Inside this single section you can use usual \LaTeX{} commands.

\nocite{lamport:latex}

\acknowledgement{Here you should mention the names of grants
                 which funded the work on this article. As defined
                 by the particular grant.}

\bibliography{common}
\bibliographystyle{iitsrc}
\end{document}